Loop Invariant Code Motion in PyPy's Tracing JIT

### 2011-06-13 talk/iwtc11/paper.tex
pure operations

\authorinfo{Hakan Ardo XXX}
           {Affiliation1}
           {hakan@debian.org}
\authorinfo{Carl Friedrich Bolz}
           {Heinrich-Heine-Universität Düsseldorf}
           {cfbolz@gmx.de}
\authorinfo{Maciej Fijałkowski}
           {Affiliation2}
           {fijall@gmail.com}

@@

In general what is needed is for the heap optimizer is to keep track of
which variables from the preamble it reuses in the peeled loop.
It has to construct a vector of such variables $H$ which
can be used to update the input and jump arguments,
\begin{equation}
  \hat J = \left(J_1, J_2, \cdots, J_{|J|}, H_1, H_2, \cdots, H_{|H}\right)
\end{equation}
\begin{equation}
  \hat K = \left(K_1, K_2, \cdots, K_{|J|}, m(H_1), m(H_2), \cdots, m(H_{|H})\right)
  .
\end{equation}
In the optimized trace $I$ is replaced by $\hat I$ and $K$ by $\hat
K$. The trace from Figure~\ref{fig:unopt-trace} will be optimized to:

@@

\subsection{Pure operation reusage}
XXX

>>>>

\authorinfo{Hakan Ardo XXX}
           {Centre for Mathematical Sciences, Lund University}
           {hakan@debian.org}
\authorinfo{Carl Friedrich Bolz}
           {Heinrich-Heine-Universität Düsseldorf}
           {cfbolz@gmx.de}
\authorinfo{Maciej Fijałkowski}
           {Affiliation2}
           {fijall@gmail.com}

@@

In general what is needed is for the heap optimizer is to keep track of
which variables from the preamble it reuses in the peeled loop.
It has to construct a vector of such variables $H$ which
can be used to update the input and jump arguments using
\begin{equation}
  \hat J = \left(J_1, J_2, \cdots, J_{|J|}, H_1, H_2, \cdots, H_{|H}\right)
  \label{eq:heap-inputargs}
\end{equation}
and
\begin{equation}
  \hat K = \left(K_1, K_2, \cdots, K_{|J|}, m(H_1), m(H_2), \cdots, m(H_{|H})\right)
  .
  \label{eq:heap-jumpargs}
\end{equation}
In the optimized trace $I$ is replaced by $\hat I$ and $K$ by $\hat
K$. The trace from Figure~\ref{fig:unopt-trace} will be optimized to:

@@

\subsection{Pure operation reusage}
If a pure operation appears more than once in the trace with same input
arguments, it only needs be executed the first time and then the result
can be reused for all other appearances. When that is combined with loop
peeling, the single execution of the operation is placed in the
preamble. That is, loop invariant pure operations are moved out of the
loop. The interactions here are the same as in the previous
section. That is, a vector, $H$, of variables produced in the preamble
and used in the peeled loop needs to be constructed. Then the jump and
input arguments are updated according to
Equation~\ref{eq:heap-inputargs} and Equation~\ref{eq:heap-jumpargs}.
